```{r}
library(Seurat)
library(magrittr)
library(dplyr)
library(ggplot2)
```

# load integrated seurat object

```{r}
obj.integrated = readRDS(
  file = file.path(
    "..",
    "..",
    "analysis_output",
    "Alistair_Utah_st_integrate",
    "visium_alistair_utah_integrate_diet.rds"
  )
)
```

color palette for seurat cluster
```{r}
cluster_color_palette = setNames(
  c(RColorBrewer::brewer.pal(n = 11, name = "Set3"), "#777777"),
  as.character(0:11)
)
```

# load cOV results

```{r}
concensusOV_results = readRDS(file = file.path("..", "..", "analysis_output", "cOV_output.rds"))
```

Extract the classification labels 
```{r}
pb_subtype_classification = concensusOV_results %>% select(sample, method, classification) %>% 
  distinct(.keep_all = FALSE) %>% 
  tidyr::pivot_wider(id_cols = sample, names_from = method, values_from = classification, names_glue = "{method}_subtype")
match_idx = match(obj.integrated$sample, pb_subtype_classification$sample)

obj.integrated@meta.data[["cOV.subtype"]] = pb_subtype_classification$consensusOV_subtype[match_idx]
obj.integrated@meta.data[["Konency.subtype"]] = pb_subtype_classification$Konecny_subtype[match_idx]
obj.integrated@meta.data[["Verhaak.subtype"]] = pb_subtype_classification$Verhaak_subtype[match_idx]
```

# Load CARD deconvolution results
```{r}
annotation_dict = read_csv(file.path("..", 
                                     "..",
                                     "data", 
                                     "scVI_integration_output", 
                                     "annotation_dict.csv"),
                           col_names = F)

deconv.props.list.CARD = readRDS(
  file = file.path(
    "..",
    "..",
    "deconvolution",
    "intermediate_output",
    "CARD",
    "props.rds"
  )
)

# merge outputs 
card.full.out.merge = NULL
for (n in names(deconv.props.list.CARD)) {
  abbr = strsplit(n, "_")[[1]][2]
  s = strsplit(n, "_")[[1]][1]
  
  df = deconv.props.list.CARD[[n]]
  spot_ids = paste0(abbr, "_" , rownames(df))
  rownames(df) = spot_ids 
  df = as_tibble(df)
  df = df %>% 
    dplyr::mutate(spot_id = spot_ids,
                  sample = s)
  
  card.full.out.merge = rbind(card.full.out.merge,
                              df)
}

# convert into long df format for plotting 
deconv_df = obj.integrated@meta.data %>% 
  dplyr::mutate(spot_id = paste0(sample, "_", barcode)) %>% 
  dplyr::select(sample, seurat_clusters, spot_id, cOV.subtype) %>% 
  dplyr::left_join(card.full.out.merge %>% dplyr::select(-c(sample)), by = "spot_id") %>% 
  dplyr::filter(!is.na(sample),
                !is.na(cOV.subtype)) %>%
  tidyr::pivot_longer(cols = -c(sample, seurat_clusters, spot_id, cOV.subtype))
deconv_df
```

# Load cell types names present in the deconvolution results and specify a order
```{r}
annotated_celltypes = unique(read.csv(file.path("..", "..", "data", "scVI_integration_output", "annotation_dict.csv"), header = F)$V2)

annotated_celltypes_order = c(annotated_celltypes[grepl("^Epithelial", annotated_celltypes)],
                              annotated_celltypes[grepl("^Mesothelial", annotated_celltypes)],
                              annotated_celltypes[grepl("^Endothelial", annotated_celltypes)],
                              annotated_celltypes[grepl("^Macrophage", annotated_celltypes)],
                              annotated_celltypes[grepl("^Monocyte", annotated_celltypes)],
                              annotated_celltypes[grepl("^pDC", annotated_celltypes)],
                              annotated_celltypes[grepl("^Plasmablast", annotated_celltypes)],
                              annotated_celltypes[grepl("^T", annotated_celltypes)],
                              annotated_celltypes[grepl("^Fibroblast", annotated_celltypes)],
                              annotated_celltypes[grepl("^Myofibroblast", annotated_celltypes)])
```

# Plot seurat integration UMAP
```{r, fig.width=6, fig.height=4}
panel.b = DimPlot(obj.integrated, 
        group.by = "seurat_clusters",
        label = T) + 
  scale_color_manual(values = cluster_color_palette) + 
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "UMAP 1", y = "UMAP 2")
panel.b

ggsave('main_fig_2_panel_b.png', 
       plot = panel.b, 
       width = 6, 
       height = 4, 
       units = "in", 
       dpi = 330, 
       bg = "white", 
       limitsize = F)
```
# Plot deconvolution classification, confidence scores across samples

```{r, fig.height=4, fig.width=6}
# specify color palette for subtype
subtype_color_palette = setNames(c("#fb746b", "#7cab04", "#0c9da9", "#c37cfb"), c("DIF", "IMR", "PRO", "MES"))

# specify a sample order
sample_sort_by_subtype = c("2B", "3C", "4D", paste0("SP", c(1,4,7,5,6,2,3,8)))

sample2source = 
  setNames(c(rep("Utah", 3), rep("Alistair", 8)), c("2B", "3C", "4D", paste0("SP", 1:8)))

method_abbreviation = setNames(c("cOV", "Konency", "Verhaak"),
                               c("consensusOV", "Konecny", "Verhaak"))

panel.a = concensusOV_results %>% 
  dplyr::mutate(source = sample2source[sample]) %>% 
  group_by(method, subtype.score) %>% 
  mutate(`scaled confidence score` = scale_0_1(score),
         Classification = classification,
         method = method_abbreviation[method],
         sample = factor(sample, levels = rev(sample_sort_by_subtype)))  %>% 
  ggplot(aes(x = subtype.score, y = sample, size = 2*`scaled confidence score`, fill = Classification)) + 
  geom_point(stroke = 0, pch = 22) + 
  facet_grid(forcats::fct_rev(source) ~ method, drop = T, scales = "free", space = "free", shrink = T) + 
  labs(x = "cOV Confidence score category", y = "Sample") + 
  scale_size_continuous(name = stringr::str_wrap(" Confidence score (scaled)", width = 18)) + 
  scale_fill_manual(name = "cOV subtype", values = subtype_color_palette) + 
  theme_minimal() + 
  theme(
        axis.text.x = element_text(angle = 90),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  guides(fill = guide_legend(override.aes = list(size = 5)))

panel.a

ggsave('main_fig_2_panel_a.png', 
       plot = panel.a, 
       width = 6, 
       height = 6, 
       units = "in", 
       dpi = 330, 
       bg = "white", 
       limitsize = F)
```
# plot sample-wise distribution of integration clustering
```{r}
panel.c = obj.integrated@meta.data %>% 
  dplyr::select(sample, seurat_clusters, barcode, cOV.subtype)%>% 
  dplyr::mutate(count = 1,
                source = factor(sample2source[sample], levels = c("Utah", "Alistair")),
                sample = factor(sample, levels = rev(sample_sort_by_subtype))) %>% 
  ggplot(aes(fill=seurat_clusters, y=sample, x=count)) + 
  geom_bar(position="fill", stat="identity", width = 0.7) + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank()) + 
  labs(y = "Samples", x = "Proportion") + 
  scale_fill_manual(values = cluster_color_palette, name = "Seurat Cluster") + 
  facet_grid(cOV.subtype ~., drop = T, scales = "free", space = "free", shrink = T) +
  theme_minimal() + 
  theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  guides(fill = guide_legend(override.aes = list(size = 5)))
panel.c

ggsave('main_fig_2_panel_c.png', 
       plot = panel.c, 
       width = 6, 
       height = 3, 
       units = "in", 
       dpi = 330, 
       bg = "white", 
       limitsize = F)
```

```{r}
# omit plotting 2 cell types with very low proportions
features = annotated_celltypes_order[!annotated_celltypes_order %in% c("T Lymphocyte(IGNG2)", "Monocyte(FCN1)")]

panel.d = deconv_df %>%
  dplyr::filter(name %in% features) %>%
  dplyr::mutate(name = factor(name, levels = features),
                subtype = factor(cOV.subtype, levels = c("DIF", "MES", "IMR"))) %>%
  ggplot(aes(y = subtype, x = value, fill = subtype)) + 
  geom_violin(scale = "width") +
  facet_wrap(name ~ ., ncol = 4, drop = T) +
  theme_bw()  + 
  scale_fill_manual(name = "cOV pseudobulk-subtype", values = subtype_color_palette, drop=T) +
  labs(x = "RCTD proprotion", y = "Subtype") + 
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) +
  xlim(0, 1) + 
  scale_y_discrete(drop=T)
panel.d

ggsave('main_fig_2_panel_d.png', 
       plot = panel.d, 
       width = 8, 
       height = 4, 
       units = "in", 
       dpi = 330, 
       bg = "white", 
       limitsize = F)
```

```{r}
cluster_order = as.character(c(0, 3, 4, 5, 6, 1, 2, 7))
panel.e = deconv_df %>%
  dplyr::filter(name %in% features) %>%
  dplyr::mutate(name = factor(name,levels = features),
                seurat_clusters = factor(seurat_clusters, levels = cluster_order)) %>%
  ggplot(aes(y = seurat_clusters, x = value, fill = seurat_clusters)) + 
  geom_violin(scale = "width") +
  facet_wrap(name ~ ., ncol = 4) +
  theme_bw() +
  scale_fill_manual(values = cluster_color_palette, name = "Seurat Cluster") +
  labs(x = "RCTD proprotion", y = "Seurat cluster") +
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  xlim(0, 1)

panel.e

ggsave('main_fig_2_panel_e.png', 
       plot = panel.e, 
       width = 8, 
       height = 4, 
       units = "in", 
       dpi = 330, 
       bg = "white", 
       limitsize = F)
```








