
```{r}
library(Seurat)
library(magrittr)
library(dplyr)
library(ggplot2)
```

# load integrated seurat object

```{r}
obj.integrated = readRDS(
  file = file.path(
    "..",
    "..",
    "analysis_output",
    "Alistair_Utah_st_integrate",
    "visium_alistair_utah_integrate_diet.rds"
  )
)
```

color palette for seurat cluster
```{r}
cluster_color_palette = setNames(
  c(RColorBrewer::brewer.pal(n = 11, name = "Set3"), "#777777"),
  as.character(0:11)
)
```

# load cOV results

```{r}
concensusOV_results = readRDS(file = file.path("..", "..", "analysis_output", "cOV_output.rds"))
```

Extract the classification labels 
```{r}
pb_subtype_classification = concensusOV_results %>% select(sample, method, classification) %>% 
  distinct(.keep_all = FALSE) %>% 
  tidyr::pivot_wider(id_cols = sample, names_from = method, values_from = classification, names_glue = "{method}_subtype")
match_idx = match(obj.integrated$sample, pb_subtype_classification$sample)

obj.integrated@meta.data[["cOV.subtype"]] = pb_subtype_classification$consensusOV_subtype[match_idx]
obj.integrated@meta.data[["Konency.subtype"]] = pb_subtype_classification$Konecny_subtype[match_idx]
obj.integrated@meta.data[["Verhaak.subtype"]] = pb_subtype_classification$Verhaak_subtype[match_idx]
```

# Load CARD deconvolution results
```{r}
card.full.out.list = list()
card.full.out.merge = data.frame()
spot_ids = NULL
for (i in 1:length(sample_abbrs)) {
  sample_abbr = sample_abbrs[i]
  
  card.full.out.list[[i]] = readRDS(file = file.path(
    "..",
    "..",
    "CARD_deconvolution",
    "intermediate_output",
    "CARD_samplewise_output_old",
    paste0("CARD_scvi_ref_full_output_", sample_abbr, ".rds")
  ))
  
  spot_ids = c(spot_ids, 
               paste0(sample_abbr, "_" , rownames(card.full.out.list[[i]])))
  
  card.full.out.merge = rbind(card.full.out.merge,
                              data.frame(sample = sample_abbrs[i], card.full.out.list[[i]], row.names = rownames(card.full.out.list[[i]])))
}


card.full.out.merge = tibble(card.full.out.merge)
colnames(card.full.out.merge)[2:length(colnames(card.full.out.merge))] = colnames(card.full.out.list[[1]])
card.full.out.merge = cbind("spot_id" = spot_ids, card.full.out.merge)
```

convert into long df format for plotting 
```{r}
deconv_df = obj.integrated@meta.data %>% 
  dplyr::mutate(spot_id = paste0(sample, "_", barcode)) %>% 
  dplyr::select(sample, seurat_clusters, spot_id, cOV.subtype) %>% 
  dplyr::left_join(card.full.out.merge %>% dplyr::select(-c(sample)), by = "spot_id") %>% 
  dplyr::filter(!is.na(sample),
                !is.na(cOV.subtype)) %>%
  tidyr::pivot_longer(cols = -c(sample, seurat_clusters, spot_id, cOV.subtype))
deconv_df
```
# Load cell types names present in the deconvolution results and specify a order
```{r}
annotated_celltypes = unique(read.csv(file.path("..", "..", "data", "scVI_output", "annotation_dict.csv"), header = F)$V2)
annotated_celltypes_order = c(annotated_celltypes[grepl("^Epithelial", annotated_celltypes)],
                              annotated_celltypes[grepl("^Endothelial", annotated_celltypes)],
                              annotated_celltypes[grepl("^Macrophage", annotated_celltypes)],
                              annotated_celltypes[grepl("^Plasmablast", annotated_celltypes)],
                              annotated_celltypes[grepl("^T", annotated_celltypes)],
                              annotated_celltypes[grepl("^Fibroblast", annotated_celltypes)],
                              annotated_celltypes[grepl("^Myofibroblast", annotated_celltypes)])
```

# Plot seurat integration UMAP
```{r, fig.width=6, fig.height=4}
panel.b = DimPlot(obj.integrated, 
        group.by = "seurat_clusters",
        label = T) + 
  scale_color_manual(values = cluster_color_palette) + 
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "UMAP 1", y = "UMAP 2")
panel.b
```
# Plot deconvolution classification, confidence scores across samples

```{r, fig.height=4, fig.width=6}
# specify color palette for subtype
subtype_color_palette = setNames(c("#fb746b", "#7cab04", "#0c9da9", "#c37cfb"), c("DIF", "IMR", "PRO", "MES"))

# specify a sample order
sample_sort_by_subtype = c("2B", "3C", "4D", paste0("SP", c(1,4,7,5,6,2,3,8)))

sample2source = 
  setNames(c(rep("Utah", 3), rep("Alistair", 8)), c("2B", "3C", "4D", paste0("SP", 1:8)))

panel.a = results %>% 
  dplyr::mutate(source = sample2source[sample]) %>% 
  group_by(method, subtype.score) %>% 
  mutate(`scaled confidence score` = scale_0_1(score),
         Classification = classification,
         method = method_abbreviation[method],
         sample = factor(sample, levels = rev(sample_sort_by_subtype)))  %>% 
  ggplot(aes(x = subtype.score, y = sample, size = 2*`scaled confidence score`, fill = Classification)) + 
  geom_point(stroke = 0, pch = 22) + 
  facet_grid(forcats::fct_rev(source) ~ method, drop = T, scales = "free", space = "free", shrink = T) + 
  labs(x = "cOV Confidence score category", y = "Sample") + 
  scale_size_continuous(name = stringr::str_wrap(" Confidence score (scaled)", width = 18)) + 
  scale_fill_manual(name = "cOV subtype", values = subtype_color_palette) + 
  theme_minimal() + 
  theme(
        axis.text.x = element_text(angle = 90),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  guides(fill = guide_legend(override.aes = list(size = 5)))

panel.a
```
# plot sample-wise distribution of integration clustering
```{r}
panel.c = obj.integrated@meta.data %>% 
  dplyr::select(sample, seurat_clusters, barcode, cOV.subtype)%>% 
  dplyr::mutate(count = 1,
                source = factor(sample2source[sample], levels = c("Utah", "Alistair")),
                sample = factor(sample, levels = rev(sample_sort_by_subtype))) %>% 
  ggplot(aes(fill=seurat_clusters, y=sample, x=count)) + 
  geom_bar(position="fill", stat="identity", width = 0.7) + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank()) + 
  labs(y = "Samples", x = "Proportion") + 
  scale_fill_manual(values = cluster_color_palette, name = "Seurat Cluster") + 
  facet_grid(cOV.subtype ~., drop = T, scales = "free", space = "free", shrink = T) +
  theme_minimal() + 
  theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  guides(fill = guide_legend(override.aes = list(size = 5)))
panel.c
```

```{r}
features = c("Epithelial cell(WFDC2)", "Epithelial cell(KRT14)", "Macrophage(CD68)", "Plasmablast", 
  "Fibroblast(COL1A1)", "Fibroblast(RAMP1)", "Fibroblast(DPT)", "Myofibroblast(PLN;ACTA2)")

panel.d = deconv_df %>%
  dplyr::filter(name %in% features) %>%
  dplyr::mutate(name = factor(name, levels = features),
                subtype = factor(cOV.subtype, levels = c("DIF", "MES", "IMR"))) %>%
  ggplot(aes(y = subtype, x = value, fill = subtype)) + 
  geom_boxplot(outlier.size = 0.25) + 
  facet_wrap(name ~ ., ncol = 4, drop = T) +
  theme_bw()  + 
  scale_fill_manual(name = "cOV pseudobulk-subtype", values = subtype_color_palette, drop=T) +
  labs(x = "CARD deconvolution proprotion estimates", y = "Subtype") + 
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) +
  xlim(0, 1) + 
  scale_y_discrete(drop=T)
panel.d
```
```{r}
panel.e = deconv_df %>%
  dplyr::filter(name %in% features) %>%
  dplyr::mutate(name = factor(name,levels = features),
                seurat_clusters = factor(seurat_clusters, levels = cluster_order)) %>%
  ggplot(aes(y = seurat_clusters, x = value, fill = seurat_clusters)) + 
  geom_boxplot(outlier.size = 0.25) + 
  facet_wrap(name ~ ., ncol = 4) +
  theme_bw() +
  scale_fill_manual(values = cluster_color_palette, name = "Seurat Cluster") +
  labs(x = "CARD deconvolution proprotion estimates", y = "Seurat cluster") +
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) + 
  xlim(0, 1)

panel.e
```
```{r}
library(patchwork)
```


```{r, fig.height=10, fig.width=14}
patch = (((panel.a | plot_spacer() | (panel.b / panel.c)) + 
    plot_layout(widths = unit(c(4, 0.5, 4), c("in", "in", "in"))))/ 
  (panel.d | panel.e)) +
  plot_layout(heights = unit(c(4, 3), c("in", "in"))) +
  plot_annotation(tag_levels = "A",
                  tag_suffix = ".")
patch
```

```{r}
ggsave(
  filename = "main_figure_deconv_subtype.png",
  plot = patch,
  path = ".",
  width = 14,
  height = 10,
  units = "in",
  dpi = 330,
  limitsize = F,
  bg = "white",
)
```

```{r}

```

```{r}

```

```{r}

```







